Make any stupid plot you want

### 2021-02-18 draft/sections/next.tex
new conclusion, make whatever plot you want

This work contributes a mathematical description of the transformation from data to visual representation. Combining Butler's fiber bundle model of data with Spivaks formalism of data schemas provides a way of decoupling topology from variability such that the model can support a very large variety of datasets, including discrete relational tables, multivariate high resolution spatio temporal datasets, and complex networks. Modeling the graphic as a fiber bundle provides a way to separate the target display space from the topology of the graphic. By decomposing the mapping from data to visual representation as encoding \vchannel, assembly \vmark, and a mapping between data and graphic topologies \vindex and formalizing what equivariance each stage needs to preserved, this work derives constraints that visualization library authors could embed in their code to guarantee visualizations that are equivariant transforms of the input data.

This work generalizes previous research constraining visual encodings from data components to graphic components as equivariant maps to components that are N-dimensional. Furthermore, it precisely defines the glyph as the visual element constructed from data on a simplex in a simplacial complex where the simplex is discrete or continuous. This is a restatement of for example Bertin's definition of a line that encapsulates all points on the continuous line.  By modeling the data topology along with its variablity, this work also provides  a generalization of topology preservation as a deformation retraction from graphic space to data space. By using a functional paradigm, we can deconstruct the graphic to the glyph associated with each data point or even to pieces of a glyph; therefore the renderer has full  flexibility in how to to generate the image, while the graphic to data topology maps provide a way to keep track of which part of the image belongs to which data point.

The toy prototype built using this model validates that is usable for a general purpose visualization tool since it can be iteratively integrated into the existing architecture rather than starting from scratch/ Factoring out glyph formation into assembly functions allows for much more clarity in how the glyphs differ. This prototype demonstrates that this framework can generate the fundemental marks, point (scatter plot), line (line chart), and area (bar chart). Furthermore, the grouped and stacked bar examples demonstrate that this model supports composition of glyphs into more complex graphics. These composite examples also rely on the fiber bundles section base book keeping to keep track of which components contribute to the attributes of the glyph. Implementing this example using a Pandas dataframe demonstrates the ease of incorporating existing widely used data containers rather than requiring users to conform to one stands.

@@

So far this model has only been worked out for a single data set tied to a primitive mark, but it should be extensible to compositing datasets and complex glyphs. The examples and prototype have so far only been implemented for the static 2D case, but nothing in the math limits to 2D and expansion to the animated case should be possible because the model is formalized in terms of the sheaf. While this model supports equivariance of figurative glyphs generated form parameters of the data\cite{beckfeathers2014,byrneFigurativeFramesCritical2017}, it currently does not have a way to evaluate the semantic accuracy of the figurative representation. This model also does not currently factor in effectiveness, but potentially effectiveness criteria could be incorporated into a scheme for assigning encoders and assembling glyphs. Also, even though the model is designed to be backend independent, it has only really been tested against the AGG backend. It is especially unknown how this framework interfaces with high performance rendering libraries such as openGL\cite{CarsonOpenGL1997}.  Because this model has been limited to the graphic design space, it does not address the critical task of laying out the graphics in the image

@@

\section{Conclusion}
Despite the alternative formalism to Mackinlay's APT framework, it is indebted to APT for providing a guideline for issues the model needs to address. As with APT, this works aims to develop a framework on which to build visualization tools. Unlike APT, our model is geared towards balancing the needs of the the scientific and information visualization communities; therefore we present a framework that can describe both how the components of the data are encoded and how the continuity is preserved in the graphic. We hope that this framework can distill the constraints on the data to graphic mapping such that it can be used to improve the architecture of a heavily used visualization tool and allow developers to more easily build domain specific tools. 
\end{document}

>>>>

This work contributes a mathematical description of the mapping \vartist\ from data to visual representation. Combining Butler's proposal of a fiber bundle model of visualization data with Spivak's formalism of schema lets this mode; support a variety of datasets, including  discrete relational tables,, multivariate high resolution spatio temporal datasets, and complex networks. Decomposing the artist into encoding \vchannel, assembly \vmark, and reindexing \vindex\ provides the specifications for producing visualization where the structure and properties match those of the input data. These specifications are that the graphic must have continuity equivalent to the data, and that the visual characteristics of the graphics are equivariant to their corresponding components under monoid actions. This model defines these constraints on the transformation function such that they are not specific to any one type of encoding or visual characteristic. Encoding the graphic space as a fiber bundle provides a structure rich abstraction of the target graphical design in the target display space.

The toy prototype built using this model validates that is usable for a general purpose visualization tool since it can be iteratively integrated into the existing architecture rather than starting from scratch. Factoring out glyph formation into assembly functions allows for much more clarity in how the glyphs differ. This prototype demonstrates that this framework can generate the fundemental marks: point (scatter plot), line (line chart), and area (bar chart). Furthermore, the grouped and stacked bar examples demonstrate that this model supports composition of glyphs into more complex graphics. These composite examples also rely on the fiber bundles section base book keeping to keep track of which components contribute to the attributes of the glyph. Implementing this example using a Pandas dataframe demonstrates the ease of incorporating existing widely used data containers rather than requiring users to conform to one stands.

@@

So far this model has only been worked out for a single data set tied to a primitive mark, but it should be extensible to compositing datasets and complex glyphs. The examples and prototype have so far only been implemented for the static 2D case, but nothing in the math limits to 2D and expansion to the animated case should be possible because the model is formalized in terms of the sheaf. While this model supports equivariance of figurative glyphs generated form parameters of the data\cite{beckfeathers2014,byrneFigurativeFramesCritical2017}, it currently does not have a way to evaluate the semantic accuracy of the figurative representation. Effectiveness is out of scope for this model because it is not part of the structure being preserved, but potentially a developer building a domain specific library with this model could implement effectiveness criteria in the artists. Also, even though the model is designed to be backend independent, it has only really been tested against the AGG backend. It is especially unknown how this framework interfaces with high performance rendering libraries such as openGL\cite{CarsonOpenGL1997}.  Because this model has been limited to the graphic design space, it does not address the critical task of laying out the graphics in the image

@@

\section{Conclusion}
An unoffical philosophy of Matplotlib is to support making whatever kinds of plots a user may want, even if they seem nonsensical to the development team. The topological framework described in this work provides a way to facilitate this graph creation in a rigorous manner; any artist that meets the equivariance criteria described in this work by definition generates a graphic representation that matches the structure of the data being represented. We leave it to domain specialists to define the structure they need to preserve and the maps they want to make, and hopefully make the process easier by untangling these components into seperate constrained maps and providing a fairly general data and display model. 
\end{document}
